---
title: "R Script using PhreeqC to Compute Solubility Curve"
output:
  html_document:
    df_print: paged
---

# Load Libraries
ggplot2 is used for graphs, dplyr might be unused
RODBC is used for database and sql--can be used to obtain data from M-Inference
RDCOMClient is used to run PhreeqC
strinr is used to formulate the input to PhreeqC

```{r}
library(ggplot2)
library(dplyr)
library(RODBC)
library(RDCOMClient)
library(stringr)
```

# Function to run PhreeqC 
This function runs a PhreeqC code passed using the input_string. A PhreeqC database must also be specified. The output is a dataframe. PhreeqC output is controlled by the SELECTED_OUTPUT block in PhreeqC. The computer running this will need to have IPhreeqc installed.

```{r}
selected_array = function(db_path, input_string) {
  dbase = COMCreate('IPhreeqcCOM.Object')
  dbase$LoadDatabase(db_path)
  dbase[["DumpStringOn"]] = TRUE
  dbase$RunString(input_string)
  selected_output = dbase$GetSelectedOutputArray()
  dump_string = dbase$GetDumpString()
  df = data.frame(sapply(selected_output, "[", 2:length(selected_output[[1]])))
  colnames(df) = sapply(selected_output, "[", 1)
  return(list(df, dump_string))
}
```

# Use a for loop over pH values and compute Al solubility
Assume Gibbsite is the solubility limiting phase. Use wateq4f for the database.

```{r}
# Specify database
db_path = r"(C:\phreeqc\database\WATEQ4F.DAT)"
# Create range of pH values in df
df = data.frame(pH = seq(from = 4, to = 10, length.out = 61))
# Add a column for Al_sol results to df
df['Al_sol'] = NA
# loop over pH values and perform PhreeqC calculation
for (i in 1:nrow(df)){
  # Get i-th pH value
  pH = df$pH[i]
  # Populate an PhreeqC input string with pH value
  input_string = str_interp("
  SOLUTION 1
    units mol/kgw
    pH ${pH}
    Al 1e-4 Gibbsite 0
  SELECTED_OUTPUT
    -high_precision	true
    -reset false
    -totals Al
  ")
  # Execute PhreeqC computation. tryCatch returns FALSE if PhreeqC computation fails
  temp_result = tryCatch({
            selected_array(db_path, input_string)[[1]]
            }, error = function(e) {FALSE})
  # Record result if PhreeqC was successful
  if (temp_result[1] != FALSE) {
    df$Al_sol[i] = temp_result$`Al(mol/kgw)`}
}
```

# Plot results

```{r}
ggplot(data = df, 
       mapping = aes(x = pH, y = Al_sol)) + 
  geom_line() +
  scale_y_continuous (trans='log10') +
  ylab("[Al] (M)")
```

